
---
title: "Exercise Set 1"
author: "T. Evgeniou"
output: html_document
---



```{r echo=TRUE, eval=FALSE, tidy=TRUE}
getwd()

setwd("Exercises/Exerciseset1/")

list.files()
```



### Exercise Data

We download daily prices (open, high, low, close, and adjusted close) and volume data of publicly traded companies and markets from the web (e.g. Yahoo! or Google, etc). This is done by sourcing the file data.R as well as some helper functions in herpersSet1.R which also installs a number of R libraries (hence the first time you run this code you will see a lot of red color text indicating the *download* and *installation* process):

```{r eval = TRUE, echo=TRUE, error = FALSE, warning=FALSE,message=FALSE,results='asis'}
list.=source("helpersSet1.R")
source("dataSet1.R")
```



#### Optional Question 

1. Can you find some interesting finance related R package or github repository? 
<br>
"Found nothing worth sharing."
<br>
<br>

<hr>
<hr>

### Part I: Statistics of S&P Daily Returns

We have `r nrow(StockReturns)` days of data, starting from `r rownames(StockReturns)[1]` until `r tail(rownames(StockReturns),1)`.  Here are some basic statistics about the S&P returns:

1. The cumulative returns of the S&P index during this period is `r round(100*sum(StockReturns[,1]),1)`%.
2. The average daily returns of the S&P index during this period is `r round(100*mean(StockReturns[,1]),3)`%;
2. The standard deviation of the daily returns of the S&P index during this period is `r round(100*sd(StockReturns[,1]),3)`%;

Here are returns of the S&P in this period (note the use of the helper function pnl_plot - defined in file helpersSet1.R):

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
SPY = StockReturns[,"SPY"]
pnl_plot(SPY)
```

#### Questions

1. Notice that the code also downloads the returns of Apple during the same period. Can you explain where this is done in the code (including the .R files used)?
2. What are the cumulative, average daily returns, and the standard deviation of the daily returns of Apple in the same period? 
3. *(Extra points)* What if we want to also see the returns of another company, say Yahoo!, in the same period? Can you get that data and report the statistics for Yahoo!'s stock, too? 
<br>
<br>
<b>Answer:
<br>
<b> 1. In R-file "dataSet1.R", the line of code "mytickers = c("SPY", "AAPL")" tells R to download SPY as well as Apple. </b>
<br>
<b>2. For apple, cumulative returns, average daily returns, and the standard deviation of daily returns look the following way: 
<br>
</b>
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
AAPL = StockReturns[,"AAPL"]
pnl_plot(AAPL)
```
<br>

<b>
- Cumulative returns: `r round(100*sum(StockReturns[,"AAPL"]),1)`%

<br>
- Average daily returns: `r round(100*mean(StockReturns[,"AAPL"]),3)`%

<br>
- Standard deviation of the daily returns: `r round(100*sd(StockReturns[,"AAPL"]),3)`%

<br>
3. In order to add YAHOO! to the dataset, we have to add the YAHOO!-ticker to the dataset using the code mytickers = c("SPY", "AAPL","YHOO"). Once, YAHOO! is in the dataset, we can easily do the same calculations that we did for Apple, giving:


</b>
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
YHOO = StockReturns[,"YHOO"]
pnl_plot(YHOO)
```
<br>


<b>
- Cumulative returns: `r round(100*sum(StockReturns[,"YHOO"]),1)`%

<br>
- Average daily returns: `r round(100*mean(StockReturns[,"YHOO"]),3)`%

<br>
- Standard deviation of the daily returns: `r round(100*sd(StockReturns[,"YHOO"]),3)`%



</b>
<br>
<br>
<br>
<br>

<hr>
<hr>

### Part II: Simple Matrix Manipulations
</b>
For this part of the exercise we will do some basic manipulations of the data. First note that the data are in a so-called matrix format. If you run these commands in RStudio (use help to find out what they do) you will see how matrices work: 

```{r eval = FALSE, echo=TRUE}
class(StockReturns)
dim(StockReturns)
nrow(StockReturns)
ncol(StockReturns)
StockReturns[1:4,]
head(StockReturns,5)
tail(StockReturns,5) 
```

We will now use an R function for matrices that is extremely useful for analyzing data. It is called *apply*. Check it out using help in R. 

For example, we can now quickly estimate the average returns of S&P and Apple (of course this can be done manually, too, but what if we had 500 stocks - e.g. a matrix with 500 columns?) and plot the returns of that 50-50 on S&P and Apple portfolio:

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig=TRUE}
portfolio = apply(StockReturns[,c("SPY","AAPL")],1,mean)
names(portfolio) <- rownames(StockReturns)
pnl_plot(portfolio)
```


We can also transpose the matrix of returns to create a new "horizontal" matrix. Let's call this matrix (variable name) transposedData. We can do so using this command:  `transposedData = t(StockReturns)`.

#### Questions

1. What R commands can you use to get the number of rows and number of columns of the new matrix called transposedData?
2. Based on the help for the R function *apply* (`help(apply)`), can you create again the portfolio of S&P and Apple and plot the returns in a new figure below?
<br>
<br>
<b> 
Answer:<br>
<br>
<b> 1. The number of rows and columns of matrix 'TransposedData' can be obtained using the functions ncol() and nrow().</b>
<br>
<b> 2. Using the functions "transposedData = tStockReturns[,c("SPY","AAPL")])", "Portfolio <- apply(transposedData,2,mean)", "names(Portfolio)<-rownames(StockReturns[,c("SPY","AAPL")])" and "pnl_plot(portfolio)", one can obtain the following view on the returns of the portfolio SPY-AAPL:
<br>
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig=TRUE}
transposedData = t(StockReturns[,c("SPY","AAPL")])
Portfolio <- apply(transposedData,2,mean)
names(Portfolio)<-rownames(StockReturns[,c("SPY","AAPL")])
pnl_plot(Portfolio)
```
</b>
<br>
<br>
<br>

<hr>
<hr>

### Part III: Reproducibility and Customization
</b> 
This is an important step and will get you to think about the overall process once again. 

#### Questions

1. We want to re-do all this analysis with data since 2001-01-01: what change do we need to make in the code (hint: all you need to change is one line - exactly 1 number! - in data.R file), and how can you get the new exercise set with the data since 2001-01-01? 
2. *(Extra Exercise)*  Can you get the returns of a few companies and plot the returns of an equal weighted portfolio with those companies during some period you select? 
<br>
<br>
<br>
<b> Answer: <br>

<b> 1. If we want to re-do all the analysis, we have to change the start data in DataSet1.R. By changing startDate = "2001-01-01" to "2000-12-29", all charts, will be updated and will start with 2001-01-02 (December 30th/31st was a weekend, the stock market was closed on Jan 1st).
<br>
<b> 2. If we add Google (GOOG) and Tesla (TSLA) to the dataset under dataSet1.R, we can create an equally weighted portfolio of SPY, Apple, Yahoo!, Google and Tesla. The returns of this portofolio between December 6th 2012 and November 25th 2016 would have looked like this: 

</b>

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig=TRUE}
transposedNewPortfolio = t(StockReturns[c(3000:4000),])
NewPortfolio <- apply(transposedNewPortfolio,2,mean)
names(NewPortfolio)<-rownames(StockReturns[c(3000:4000),])
pnl_plot(NewPortfolio)
```

<br>
<br>
<br>
<br>

<hr>
<hr>

###  Part IV: Read/Write .CSV files
</b>
Finally, one can read and write data in .CSV files. For example, we can save the first 20 days of data for S&P and Apple in a file using the command:

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
write.csv(StockReturns[1:20,c("SPY","AAPL")], file = "twentydays.csv", row.names = TRUE, col.names = TRUE) 
```

Do not get surpsised if you see the csv file in your directories suddenly! You can then read the data from the csv file using the read.csv command. For example, this will load the data from the csv file and save it in a new variable that now is called "myData": 

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
myData <- read.csv(file = "twentydays.csv", header = TRUE, sep=";")
```

Try it!

#### Questions

1. Once you write and read the data as described above, what happens when you run this command in the console of the RStudio: `sum(myData != StockReturns[1:20,])`
2. *(Extra exercise)* What do you think will happen if you now run this command, and why:  

```{r eval = FALSE, echo=TRUE}
myData + StockReturns[1:40,]
```

<b> Answer:
<br>
1. Entering this function returns: `r sum(myData != StockReturns[1:20,])`.
<br>
2. You get an error if you try this function because myData is 20 rows of data where StockReturns is a matrix of 20X3.
<br>
<br>
<br>
<br>

<hr>
<hr>



<br>


<br>






